Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-07-20 text/paper/paper1.tex

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}
\renewcommand{\AE}{\as}
\newcommand{\GE}{\es}

@@

\svnInfo $Id$

@@

Summing up so far, three previous studies have presented diverging
pieces of evidence concerning the status of putative local
readings. \citet{GeurtsPouscoulous2009:Embedded-Implic} did not observe any truth value judgments that would indicate
local readings. On the basis of this finding they argue that these readings are simply not available. On the other
hand, C\&D and C\&S argue that the findings of G\&P may be due to a bias in their experimental design and
employed different methods intended to be sensitive enough to reveal effects of local readings: either by
a choice of the situation that better befits a given sentence (\citeauthor{CliftonDube2010:Embedded-Implic}), or by a graded
truth-value judgment (\citeauthor{ChemlaSpector2010:Experimental-Ev}). Both studies found evidence in support of local readings.

To account for all of these observations, it could be hypothesized
along with \citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} that local readings are
available, but so dispreferred that they do not affect
\emph{categorical} truth-value judgements (See, however, \citet{Crain1998} for arguments in favour of truth value judgments
as a means of detecting highly dispreferred readings).
It could then furthermore be argued, in line with traditionalism, that the reason why local
readings do not affect categorical judgements is because they are not
(strongly) truth-relevant speaker-intended meaning enrichments after
all, so that no case can be made for lexicalism or grammaticalism
\citep[c.f.][for arguments along these lines]{Tielvan-Tiel2012:Embedded-Scalar}.

The main point we would like to make in this section is that neither
interpretation is fully satisfactory because the situation is even
more complicated than suggested so far. Therefore, further empirical
testing is required, but should be carefully designed so as to
circumnavigate, as far as possible, the delicate methodological
pitfalls the subject matter so unkindly provides.

In outline our position, which we will elaborate on in the following,
is this. On the one hand, we believe with
\citet{ChemlaSpector2010:Experimental-Ev} and
\citet{Tielvan-Tiel2012:Embedded-Scalar} that properties of the
pictorial material can affect subjects' judgements. We concede,
following \citet{Tielvan-Tiel2012:Embedded-Scalar}, that a notion of
typicality may partly be responsible. But we also submit, in line with
\citet{ChemlaSpector2010:Experimental-Ev}, that typicality might not
be all of the story. On the other hand, we present suggestive evidence
based on a pilot study of ourselves (to be addressed presently) that
even categorical judgements \emph{can} be affected by pictorial
properties. We conclude that further empirical evidence is needed and
that a follow-up study should meet the following requirements: (i) it
should be a categorical truth-value judgement task after all because
only that way the traditionalist claim that any observed effects are
not truth-relevant can be suitably addressed\dn{@Fabian: would a
  reference fit here to the book(?) that discusses why true/false
  judgements are best?} and (ii) it must make sure to minimize any
putative effects of the pictorial material as much as possible.

In support of this position, we would like to briefly discuss the
results of a small pilot study of ourselves.\fn{A word of
  motivation. We report the results of this pilot mainly for
  expository reasons. A better argument that categorical judgements
  can be indicative of local readings is given by our main
  experiment. But in order to motivate the readers that they should be
  interested in this experiment, we feel it is helpful to discuss the
  mildly suggestive, but ultimately inconclusive evidence of this
  pilot. Additionally, introducing the pilot also serves another
  purpose, namely the discussion of the results for sentences like
  (\ref{bsp:EA}). These help introduce in a tangible way how seemingly
  innocuous properties of the pictorial material can influence
  subjects' truth-value judgements, and why typicality might not be
  the whole explanation for these pictorial effects.} Our pilot tried
to replicate \citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s
results, using a simple categorical truth-value judgement task on
pictures that were exactly like those in Figures
\ref{fig:AS-distinguishing-pics} and
\ref{fig:ES-distinguishing-pics}.\dn{We should expand on the method in
  an appendix} Test sentences were German cleft-constructions as in
(\ref{bsp:AS-K1}) and (\ref{bsp:ES-K1}).\fn{We concede that these
  cleft-constructions might unduly emphasize potential local readings,
  perhaps because they induces a quotative reading. (We are grateful
  to Philippe Schlenker for raising this latter possibility.) But that
  is not essential, because the main issue we are trying to address
  here is whether putative local readings \emph{could} affect
  categorical truth-value judgements.}
\begin{exe}
\ex \gll F\"{u}r \mymark{jedes} dieser Kinder gilt: es
  mag \mymark{einige} dieser Speisen. \label{bsp:AS-K1}\\
  For every {of these} children {it holds} it likes some {of
    these} dishes.\\
  For all of these children it holds that it likes some of these dishes.
\end{exe}
\begin{exe}
\ex \gll F\"{u}r \mymark{genau} \mymark{eines} dieser Kinder gilt: es
  mag \mymark{einige} dieser Speisen. \label{bsp:ES-K1}\\
  For exactly one {of these} children {it holds} it likes some {of
    these} dishes.\\
  For exactly one of these children it holds that it likes some of these dishes.
\end{exe}
Mean acceptance rates of 60 native German speakers for our target
sentences were:
\begin{center}
  \begin{tabular}{cccc}
    \multicolumn{4}{c}{\as-sentences}     \\ \cmidrule(r){1-4}
    false & literal & weak & strong 
    \\ \midrule 
    .9 & 72 & 79 & 90 
  \end{tabular}
  \qquad
  \begin{tabular}{cccc}
    \multicolumn{4}{c}{\es-sentences}    \\ \cmidrule(r){1-4} 
    false & local & literal & all
    \\ \midrule 
    .8 & 20 & 71 & 93 
  \end{tabular}
\end{center}
% \begin{center}
%   \begin{tabular}{ccccccccc}
%     \multicolumn{4}{c}{\as-sentences} &&
%     \multicolumn{4}{c}{\es-sentences}     \\ \cmidrule(r){1-4} \cmidrule(r){6-9}
%     false & literal & weak & strong && false & local & literal & all
%     \\ \midrule 
%     .9 & 72 & 79 & 90 && .8 & 20 & 71 & 93 
%   \end{tabular}
% \end{center}
Two observations are of interest here. Firstly, we observed 21\%
\emph{false}-judgements in the \as-weak condition, as well as 20\%
\emph{true}-judgements in the \es-local condition, contrary to the 0\%
in each case observed by
\citet{GeurtsPouscoulous2009:Embedded-Implic}. Secondly, the observed
difference between the \as-weak and the \as-strong conditions, on the
one had, and the \es-false and the \es-local condition, on the other
hand, both reached significance.\dn{spell out in appendix} This
suggests that local readings \emph{are} available and \emph{can} have
an effect even on categorical truth-value judgements.

% \begin{figure}[]
% \centering
% \subfloat[\as-sentences]{
% \includegraphics[width=5.5cm]{../pictures/paper/JudgmentsAllSomeBars.pdf}
% \label{fig:EABars}
% }
% \subfloat[\es-sentences]{
% \includegraphics[width=5cm]{../pictures/paper/JudgmentsExactlySomeBars.pdf}
% \label{fig:EEBars}
% }
% \caption{Means of \emph{true}-judgments in our pilot study}
% \label{fig:Bars-K1}
% \end{figure}

These observed patterns suggest that local readings \emph{are}
available and \emph{can} have an effect even on categorical
truth-value judgements.

To illustrate the potential conflation, consider a sentence like
(\ref{bsp:EA}), which is a ``global alternative'' to an \es-sentence,
with a universal quantifier where an \es-sentence has scalar
\emph{some}.
\begin{exe}
\ex \gll F\"{u}r \mymark{genau} \mymark{eines} dieser Kinder gilt: es
  mag \mymark{jede} dieser Speisen. \label{bsp:EA}\\
  For exactly one {of these} children {it holds} it likes all {of
    these} dishes.\\
  For exactly one of these children it holds that it likes all of these dishes.
\end{exe}
Sentences like (\ref{bsp:EA}) are literally true in both of the
situations in Figure~\ref{fig:EA-distinguishing-pics}, but these
situations differ in the truth value of the corresponding
\es-sentence.
%
\begin{figure}[]
  \centering

\subfloat[fig:global][global]{
  \label{fig:global-EA}

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true
          \end{tabular}};

} \qquad \qquad
\subfloat[fig:literal][literal]{
  \label{fig:literal-EA}

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & false
          \end{tabular}};

      \end{tikzpicture}
}

\caption{Distinguishing scenarios for sentence(\ref{bsp:EA})}
  \label{fig:EA-distinguishing-pics}

\end{figure}
%
Our pilot study revealed a difference in acceptability, but a curious
one. The sentence was judged \emph{true} in 90.3\% of the cases for
the situation in Figure~\ref{fig:global-EA} and in 98.3\% in situation
\ref{fig:literal-EA}. The observed difference is significant ($t =
-3.8259$, $df = 322.375$, $p < 0.001$). This is a curious finding,
because, from the point of view of implicature calculation, the only
thing that could plausibly explain this acceptability difference is
the presence of a global implicature, resulting from the negation of
the corresponding \es-sentence.\fn{Notice that (\ref{bsp:EA}) is the
  alternative to the corresponding \es-sentence needed for computing
  the putative global implicature of that \es-sentence. But since
  these sentences are logically independent, it may be assumed that
  the corresponding \es-sentence is an alternative also to sentences
  like (\ref{bsp:EA}). It may also be that \emph{being-an-alternative}
  is not a symmetric notion. Be that as it may, our argument does not
  hinge on whether (\ref{bsp:EA}) does or does not actually have a
  global implicature of the alleged kind.} But such a putative global
implicature would be true in the situation~\ref{fig:global-EA}, but
false in the situation~\ref{fig:literal-EA}. Consequently, the
observed difference in acceptability cannot be explained simply in
terms of the hypothesis of \citet{ChemlaSpector2010:Experimental-Ev}
that acceptability is proportional to the number of available true
readings. Rather some feature of the pictorial material that is
independent of the putative readings under scrutiny does affect
truth-value judgements, even categorical ones.

\citet{ChemlaSpector2010:Experimental-Ev} acknowledge the possibility
that the \emph{typicality} of a picture may affect graded truth-value
judgements. Stronger still, they show that graded judgments of
\as-sentences differ significantly for different instantiations of the
conditions in \ref{fig:AS-distinguishing-pics}, that vary in the
amount of connections without changing the assignment of truth values
to candidate readings. \citeauthor{ChemlaSpector2010:Experimental-Ev}
suggest that typicality of the pictorial material can account for
these differences, but submit that this does not explain away the
effect of the local reading. This is disputed by
\citet{Tielvan-Tiel2012:Embedded-Scalar}, who demonstrates that all of
\citeauthor{CliftonDube2010:Embedded-Implic}'s and
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s data on \as-sentences
can be explained as typicality effects, dispensing the need to appeal
to any influence of local readings. We are sympathetic towards
\citeauthor{Tielvan-Tiel2012:Embedded-Scalar}'s innovative line of
reinterpretation of the data, but note, as
\citet{ChemlaSpector2010:Experimental-Ev} already observed, that his
typicality-based explanation does not extend to \es-sentences in an
obvious way\fn{For similar reasons, appeal to typicality cannot
  explain, as far as we can see, the preference for the situation in
  Figure~\ref{fig:literal-EA} over that in Figure~\ref{fig:global-EA}
  given a sentence like (\ref{bsp:EA}).} and that it also does not
explain why we observed putative effects of typicality also in
categorical truth-value judgements in our pilot.

To conclude, the confounding effects of pictorial complexity and/or
typicality are real and affect even categorical truth-value
judgements. The available evidence about availability, preference and
nature of local readings is therefore still inconclusive. What is
needed to complement the picture is data from a design that elicits
categorical truth-value judgements that are as independent as possible
of possibly conflating properties of the pictorial material. Given the
logical dependencies of the relevant readings of \as- and
\es-sentences this seems almost impossible to achieve. As we will
argue below we do, however, believe it is.

On top of this, we would like to obtain further information about the
relative preferences among all putative readings, so as to be able to
decide between theoretical positions, as outlined in
Section~\ref{sec:theories-predictions}. Neither of the previous
studies gives us that, either because they simply have not
investigated all the relevant comparisons
(\citet{GeurtsPouscoulous2009:Embedded-Implic},
\citet{CliftonDube2010:Embedded-Implic}), or because they run risk of
conflating preferences for readings with preferences for pictures
(\citet{ChemlaSpector2010:Experimental-Ev}, our pilot study).

% Another desideratum for a further empirical testing is that a study
% aimed at revealing information about availability and preferences of
% readings be accompanied by a suitable comparable investigation into a
% phenomenon which is know to allow for several readings with a given
% known preferences structure. This way and only this way can a proper
% baseline be established to guide the interpretation of the observed
% data.

Another potential shortcoming of previous studies is that they did not
control for possible conflating effects of ``silent intonation''
(XYZ)\dn{references}. Even traditionalism maintains that local
readings are attested, but requiring a contrastive stress on the
scalar item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar}. Variation
in the interpretation of target sentences therefore could be caused by
subjects imposing different intonational patterns on the visually
presented target sentences. Ideally, we should therefore present
sentences auditorily and address directly the frequently voiced
hypothesis that local readings are triggered by contrastive
stress.\dn{relate to \citet{SchwarzClifton2008:Strengthening-o} and
  others?}

The following section presents a study designed to meet all of these
desiderata.

>>>>

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}
\renewcommand{\AE}{\as}
\newcommand{\GE}{\es}
\newcommand{\ea}{\acro{ea}}

@@

\svnInfo $Id: paper.tex 102 2012-07-04 10:19:59Z michael.franke $

@@

Summing up so far, three previous studies have presented diverging
pieces of evidence concerning the status of putative local
readings. In their picture verification experiment, \citet{GeurtsPouscoulous2009:Embedded-Implic} did not observe any truth value judgments that would indicate
local readings. On the basis of this finding they argue that these readings are simply not available. On the other
hand,  \citeauthor{CliftonDube2010:Embedded-Implic} and \citeauthor{ChemlaSpector2010:Experimental-Ev} argue that the findings of  \citeauthor{GeurtsPouscoulous2009:Embedded-Implic} may be due to a bias in their experimental design and
employed different methods intended to be sensitive enough to reveal effects of local readings: either by
a choice of the situation that better befits a given sentence (\citeauthor{CliftonDube2010:Embedded-Implic}), or by a graded
truth-value judgment (\citeauthor{ChemlaSpector2010:Experimental-Ev}). Both studies found evidence in support of local readings.
To account for the diverging results we could draw a distinction between \emph{categorical} truth-value
judgments and other more sensitive measures. Firstly, it could be hypothesized
along with \citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} that local readings are
available, but so dispreferred that they do not affect
\emph{categorical} truth-value judgements (See, however, \citet{Crain1998} for arguments in favour of truth value judgments
as a means of detecting highly dispreferred readings).
Secondly it could also be argued, in line with traditionalism, that the apparent difference in
sensitivity of the different judgment types can be explained by the stipulation that
the results reported by \citeauthor{CliftonDube2010:Embedded-Implic}
and \citeauthor{ChemlaSpector2010:Experimental-Ev} reflect something other than
(strongly) truth-relevant speaker-intended meaning enrichments after all, so that no case
can be made for lexicalism or grammaticalism on the basis of these data \citep[c.f.][for arguments along these lines]{Tielvan-Tiel2012:Embedded-Scalar}.

Note, however, that the type of judgment  is not the only crucial difference between the methods of \citeauthor{CliftonDube2010:Embedded-Implic} and \citeauthor{ChemlaSpector2010:Experimental-Ev} on the one hand and  \citet{GeurtsPouscoulous2009:Embedded-Implic} on the other. For example these studies  also differed with respect to the types of situations that were shown to the participants during the course of the experiment.  The presence or absence of competing situations may have led to skewed distributions of judgments for situations like in Figure \ref{fig:weak}. Such effects might emerge simply because the type of sentence-picture combinations shown during the experiment affects the standard of judgment but could~-- maybe more elegantly~-- also be explained by findings from \citet{Raffray2010} suggesting that exposing participants to models may prime corresponding logical forms and thereby shift reading preferences. Taking these possibilities into account, it remains unclear whether the effects reported by  \citeauthor{CliftonDube2010:Embedded-Implic} and \citeauthor{ChemlaSpector2010:Experimental-Ev} would also show up using ordinary picture verification, in the first place.

Be this as it may, we cannot ascribe the effects reported by  \citeauthor{CliftonDube2010:Embedded-Implic} and \citeauthor{ChemlaSpector2010:Experimental-Ev} to the availability of local readings with certainty. \citet{ChemlaSpector2010:Experimental-Ev} themeselves acknowledge
the possibility that the \emph{typicality} of a picture with respect to some meaning may affect graded truth-value
judgements. Strengthening this possibility, they show that graded judgments of
\as-sentences differ significantly for different instantiations of the
conditions in \ref{fig:AS-distinguishing-pics}, that vary in the
amount of connections without changing the assignment of truth values
to candidate readings. \citeauthor{ChemlaSpector2010:Experimental-Ev}
suggest that typicality of the pictorial material can account for
these differences, but submit that this does not explain away the
high acceptance of pictures like Figure \ref{fig:weak}. The latter point is disputed by
\citet{Tielvan-Tiel2012:Embedded-Scalar}, who demonstrates that a huge chunk of variance in the data of
\citeauthor{CliftonDube2010:Embedded-Implic} and \citeauthor{ChemlaSpector2010:Experimental-Ev} on
\as-sentences can be explained as typicality effects, dispensing the need to appeal
to the distribution of different readings. \citet{Tielvan-Tiel2012:Embedded-Scalar} elicited what he calls the typicality
structures associated with the quantifiers {\it all} and {\it some} and predicted the judgments of \as-sentences obtained by C\&S based
on these data. Thereby, he obtained an excelent model fit. We are sympathetic towards
\citeauthor{Tielvan-Tiel2012:Embedded-Scalar}'s innovative line of
reinterpretation of the data, but note, as
\citet{ChemlaSpector2010:Experimental-Ev} already observed, that his
typicality-based explanation does not extend to \es-sentences in an
obvious way\fn{For similar reasons, appeal to typicality cannot
  explain, as far as we can see, the preference for the situation in
  Figure~\ref{fig:literal-EA} over that in Figure\ref{fig:global-EA}
  given a sentence like (\ref{bsp:EA}).}.

To conclude, although considerable progress has been made in understanding which factors
have to be controlled for in order to asses the status of loacl readings, the evidence about availability,
preference and nature of local readings is still inconclusive. What is needed to complement
the picture are data that are sensitive enough to detect local readings even if they
are strongly dispreferred and are as independent as possible of possibly conflating properties
of the pictorial material. Given the logical dependencies of the
relevant readings of \as- and \es-sentences this seems almost
impossible to achieve. As we will argue below we do, however, believe it is.

On top of this, it would be desirable to obtain further information about
the relative preferences among all putative readings, so as to be able
to decide between theoretical positions, as outlined in
Section~\ref{sec:theories-predictions}. Neither of the previous
studies gives us that, either because they simply have not
investigated all the relevant comparisons
(\citet{GeurtsPouscoulous2009:Embedded-Implic},
\citet{CliftonDube2010:Embedded-Implic}), or because they cannot safely
derive the required cardinal information for danger of conflating
preferences for readings with preferences for pictures
(\citet{ChemlaSpector2010:Experimental-Ev}, our pilot study).

Moreover, a final omission of previous studies is that they did not
control for possible conflating effects of ``silent intonation''
(XYZ)\dn{references}. Even traditionalism maintains that local
readings are attested, but requiring a contrastive stress on the
scalar item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar}. Variation
in the interpretation of target sentences therefore could be caused by
subjects imposing different intonational patterns on the visually
presented target sentences. Ideally, we should therefore present
sentences auditorily and address directly the frequently voiced
hypothesis that local readings are triggered by contrastive
stress.\dn{relate to \citet{SchwarzClifton2008:Strengthening-o} and
  others?}

\section{Pilot Study}
To start off, we conducted  a pilot study  investigating whether the diverging evidence reviewed above can be explained by considerations on how sensitive the different kinds of measures are. This information was critical to our main experiment below. Because the main experiment employed categorial truth-value judgments out of methodological necessity, we wanted to test the sensitivity of this measure independently. In addition, the question is interesting in its own right since it has implications about the nature of the ambiguity at hand, namely whether it is due to truth-relevant meaning enrichment.

In particular, we tested whether the critical effects \citet{ChemlaSpector2010:Experimental-Ev} found in their graded truth-value judgment task do also show up in a categorial truth value judgment task. In order to test this we simply replicated \citeauthor{ChemlaSpector2010:Experimental-Ev}'s experiment 
using a categorial truth-value judgment task. If, as has been suggested, categorial truth-value judgments are not sensitive enough to detect a difference of acceptance between situations like \ref{fig:weak} and \ref{fig:strong} given \as-sentences or between \ref{fig:local} and \ref{fig:all} given \es-sentences, these should be judged identical in a truth value judgment task.

In addition to the conditions from  \citet{ChemlaSpector2010:Experimental-Ev} we also included another type of sentence. This type of sentence, \ea-sentences, contained the scalar item {\it all} embedded under {\it exactly one}. Here, a rather atypical kind of global implicature would be arrived at by replacing {\it all} with its weaker alternative {\it some}.  We tested whether this kind of implicature is detectable using a truth-value judgment task. In the discussion of the present experiment we will, however, mainly focus on the replication of  \citet{ChemlaSpector2010:Experimental-Ev}.

\subsection{Materials}
Test sentences were German cleft-constructions as in (\ref{bsp:AS-K1})$-$(\ref{bsp:EA-K1}). Sentences like (\ref{bsp:AS-K1}) are \as-sentences, (\ref{bsp:ES-K1}) are \es- and  (\ref{bsp:EA-K1}) \ea-sentences. We decided to use cleft constructions in order to fix the relative scope of the quantifiers involved. This was ensured via clause-boundedness. In total, 48 experimental items were constructed in all three conditions yielding a total of 144 sentences.

\begin{exe}
\ex \gll F\"{u}r \mymark{jedes} dieser Kinder gilt: es
  mag \mymark{einige} dieser Speisen. \label{bsp:AS-K1}\\
  For every {of these} children {it holds} it likes some {of
    these} dishes.\\
  For all of these children it holds that it likes some of these dishes.
\end{exe}
\begin{exe}
\ex \gll F\"{u}r \mymark{genau} \mymark{eines} dieser Kinder gilt: es
  mag \mymark{einige} dieser Speisen. \label{bsp:ES-K1}\\
  For exactly one {of these} children {it holds} it likes some {of
    these} dishes.\\
  For exactly one of these children it holds that it likes some of these dishes.
\ex \gll F{\"u}r \mymark{genau eines} dieser Kinder gilt: es mochte \mymark{jede} dieser Speisen.  \label{bsp:EA-K1}\\
For exactly\ one of\ these children it\ holds: it likes some of\ these dishes.\\
Exactly one of\ these children it holds that it likes some of\ these dishes.\\
\end{exe}

The sentences were paired with pictures like in Figures \ref{fig:AS-distinguishing-pics}, \ref{fig:ES-distinguishing-pics}, \label{fig:EA}. For \es- and \as-sentences  there were four types of situations just as in  the study of \citet{ChemlaSpector2010:Experimental-Ev}. Sentences in the \as-condition were paired with situations as in \ref{fig:literal-AE}, \ref{fig:false-AE}, \ref{fig:weak}, \ref{fig:strong}. Sentences in in the \es-condition were paired with \ref{fig:literal-GE},\ref{fig:literal-GE},\ref{fig:local} and \ref{fig:all}. The \ea-sentences were paired with pictures like in \ref{fig:EA}. Figure \ref{fig:EA-LIT} is only true under the literal reading ({\it literal} diagrams) while  Figure \ref{fig:EA-LITGLB} is also true with respect to the global implicature ({\it global} diagrams) denying the weaker scalar alternative containig {\it einige} (some). Diagrams like those in Figures \ref{fig:EA-FALSE1} and \ref{fig:EA-FALSE2} were not compatible with the \ea-sentences. Figure \ref{fig:EA-FALSE1} is incompatible because there are too few connections. Figure \ref{fig:EA-FALSE2} is incompatible because there are too many connections. We used two types of incompatible pictures only because of thechnical reasons.

In total, this yielded twelve conditions and 576 pictures. The conditions were distributed over twelve lists according to a latin square design ensuring that in each list every item was only present in one condition. In addition, to the target sentences, 201 (?) distractors were included in each list. Of these $n$ were intended to yield a "yes true" response, $m$ were false.

\begin{minipage}{6cm}
     \subfloat[][literal]{
    \label{fig:EA-LIT}
        \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer09.eps}
  }

  \subfloat[test][global]{
    \label{fig:EA-LITGLB}
    \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer10.eps}
  }  
   \end{minipage}
   \begin{minipage}{6cm}
     \subfloat[][false1]{
    \label{fig:EA-FALSE1}
        \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer11.eps}
  }

  \subfloat[test][false2]{
    \label{fig:EA-FALSE2}
    \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer12.eps}
  }  
   \end{minipage}
        \caption{}
\label{fig:EA}
\end{figure}

\subsection{Procedure}
The present pilot study was conducted as part of a dual task experiment employing picture verification and a sternberg task \cite{Sternberg1966}.  In some trials participants had to solve a memory task in addition to the picture verification task. In the {\it high load} condition participants had to remember six letters while performing the picture verification. In the {\it low load} condition they had to remember one letter and in the {\it no load} condition they didn't have to remember anything while they were performing the picture verification. In the {\it memory only} conditions they only had to perform a memory task (i.e. remember one or six letters) but no picture verification. All the critical items of the present experiment were presented in the {\it no load} condition. Only distractor items were presented in one of the dual task conditions.

In each trial participants were first asked to indicate readiness by pressing a button. In the {\it no load} condition, after they had pressed the button an  asterisk was flashed on the screen six times within $7.2s$. Then a sentence and a picture were presented simultaneously. Within $12s$ participants had to judge whether the sentence was true with respect to the picture. They indicated their judgment by pressing one of two buttons. After they had provided their judgment the next trial started automatically. If they did not respond within twelve seconds, the next trial was started. Judgments and reaction times were logged.\footnote{In the {\it low load} and {\it high load} condition a trial looked similar to a trial in the {\it no load} condition except for two things. Firstly, in the dual task conditions instead of asterisks letters were flashed on the screen. Secondly, after finishing the picture verification task a question appeared on the screen probing for a letter. Participants had to indicate by pressing a button whether they thought the letter had been present in the array from the beginning of the trial. Except for one difference, the {\it memory only} trials were identical to the dual task trials. In {\it memory only} trials the picture verification part was replaced with a blank screen. That is, in the middle of such trials the screen stayed blank for $5s$. }

The experiment started with a practise session consisting of $25$ trials devided into three blocks (1st block: memory task only, ten trials; 2nd block: picture verification only, five trials; 3rd block: dual task, ten trials). Then one list consisting of 249 trials devided into three blocks was presented. An experimental session took about 75 minutes. During the whole experiment no feedback was provided to the participants.

\subsection{Participants}
60 native German speakers (40 female, mean age: $24$, ranging between $20$ and $40$) from the University of T{\"u}bingen participated in the study. Participants were na\"ive to the purpose of the study. They were paid \euro{10} compensation.

\subsection{Results}
Overall participants were able to solve the task. In the distractor items the picture verification was done accurately in $88\%$ ($sd: 33\%$) of the cases. All participants were above $70\%$ and significantly better than chance. On average it took participants $4989ms$ ($sd:2176ms$) to come up with their answer in the filler trials. We will now consider results for the three sentence conditions in turn.

The mean judgments we obtained for \as-sentences are depicted in Figure \ref{fig:AEBars}. The {\it strong} diagrams were judged to be true in $90\%$ of the cases, {\it weak} diagrams received "yes" judgments in $79\%$ , {\it literal} diagrams in  $72\%$  and {\it false} in $0.9\%$ of the trials.  A logit mixed effects model with binomial link function containing a fixed effect of {\it diagram} as well as random effects of items and participants revealed a clear effect of {\it diagram}. In particular {\it false}, {\it literal} and {\it weak} diagrams all received fewer "yes" judgments than {\it strong} diargams (all $p<.001$).  Another logit mixed effects model only modelling data for {\it literal} and {\it weak} diagrams also revealed an reliable effect of {\it diagram} (p<.05). Finally, a model only considering  {\it false} and {\it literal} diagrams also revealed a significant effect of {\it diagram} (p<.001).

     \subfloat[][\as]{
    \label{fig:AEBars}
        \includegraphics[width=5cm]{../pictures/JudgmentsAllSomeBars.eps}
  }
  \subfloat[test][\es]{
    \label{fig:GEBars}
    \includegraphics[width=5cm]{../pictures/JudgmentsExactlySomeBars.eps}
  }

        \caption{Mean judgments}
\label{fig:BarsPilot}
\label{fig:EA}
\end{figure}

The responses for \es-sentences are summarized in Figure \ref{fig:GEBars}. Unsurprisingly, these sentences were accepted as truthfull description of {\it false} diagrams in only $0.8\%$ of all cases. The {\it local} diagrams received  "yes" responses in $20\%$, {\it literal} in $71\%$ and {\it all} in $93\%$ of the time. A logit mixed effects model with {\it diagram} as fixed effect as well as item and participant as random effects revealed a signficant effect of {\it diagram}. The {\it all}, {\it literal} and {\it local} diagrams were all accepted more often than the {\it false} diagram (all $p<.001$). A simpler model only considering {\it local} and {\it literal} diagrams revealed a reliable effect of {\it diagram} (p<.001) as did a model only considering {\it literal} and {\it all} diagrams (p<.001).

The mean responses for the \ea-sentences are shown in table \ref{}. The observed difference between the global and literal diagrams is
significant ($t = -3.8259$, $df = 322.375$, $p < 0.001$).

\begin{center}
  \begin{tabular}{cccc}
    \multicolumn{4}{c}{\as-sentences}  \\
    \hline
    literal & global & false1 & false2
    \\ \midrule 
    98.3 & 90.3 & 7.1 & 3.3 
  \end{tabular}
\end{center}

\subsection{Discussion}

Concerning the \as- and \es-sentences the observed distribution of judgments were rather similar to the results from \citeauthor{ChemlaSpector2010:Experimental-Ev}. Qualitatively, the only notable difference lies in the rather low acceptance of situations like \ref{fig:literal-AE} in the present experiment as compared to \citeauthor{ChemlaSpector2010:Experimental-Ev}'s original study. Crucially, similar to the findings of \citeauthor{ChemlaSpector2010:Experimental-Ev}, we found that situation \ref{fig:strong} was more often accepted than situation \ref{fig:strong} and that situation \ref{fig:local} was accepted more often than situation \ref{fig:false-GE}.  This suggests that local readings are available and are even detectable via categorial truth-value judgments.

However, we have to be careful with our interpretation of the results. Since the picture materials differd drastically in the critical comparisons, we cannot exclude pictorial effects. As reviewed above  \citeauthor{Tielvan-Tiel2012:Embedded-Scalar} did even show that with regards to the \as-sentences the differences in acceptance reported by C\&S and therefore presumably also the replicated results presented here can be explained by typicality of the pictures with respect to the sentence meaning.

That truth is not the only determining factor affecting truth-value judgments also becomes evident from the distribution of judgments we obtained for the \ea-sentences. The observed difference  between \ref{fig:EA-LIT} and \ref{fig:EA-LITGLB}  is a curious finding at first sight. Firstly, the literal reading is true in both situations. Secondly, from the point of view of implicature calculation, the only thing that could plausibly explain an acceptability difference between these two situations is the presence of a global implicature, resulting from the negation of the corresponding \es-sentence.\fn{Notice that (\ref{bsp:EA}) is the alternative to the corresponding \es-sentence needed for computing the putative global implicature. But these sentences are logically independent. So, it
  may be that the corresponding \es-sentence is an alternative also to sentences like (\ref{bsp:EA}). It may also be that being an
  alternative is not a symmetric notion. Be that as it may, our argument does not hinge on whether (\ref{bsp:EA}) does or does not
  actually have a global implicature of the alleged kind.} However, this putative global implicature would be true in the
situation \ref{fig:EA-LITGLB}, but false in the situation \ref{fig:EA-LIT}. Consequently, the observed difference
in acceptability cannot be explained simply in terms of the hypothesis
of \citet{ChemlaSpector2010:Experimental-Ev} that acceptability increases with  the number of available true readings. Rather this finding, just as  \citeauthor{Tielvan-Tiel2012:Embedded-Scalar}'s study, suggests that some
feature of the pictorial material that is independent of the putative readings under scrutiny does affect truth-value judgements, even
categorical ones.

We conclude, that  effects of pictorial complexity and/or typicality are real and affect even categorial judgments. For this reason the following experiment was set up to investigate the availability of local readings while minimizing pictorial effects such as typicality or complexity.
